Chaotic dispersal of tidal debris

### 2015-07-17 ms.tex
remove figure add figure

\todo{apw}{talk about streams, sanders/bovy stuff in actions / frequencies. streams are 1D stuctures. small deviations orthogonal to main elongation, e.g., from small-scale chaotic diffusion can matter}

An ensemble of regular orbits (e.g., tidal debris) will phase-mix due to (small) differences in the fundamental frequencies of the orbits. Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and are thus difficult to predict. For example,  ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume via the short-time chaotic diffusion processes, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.

In this work, we investigate the consequences of this short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}. We then discuss how this would affect our understanding of the coherence and density evolution of tidal streams.

@@

The MLE, $\lambda_{\rm max}$, is interpreted as a rate that quantifies the exponential divergence of infinitesimally close chaotic orbits. It is therefore useful to consider the corresponding $e$-folding time by inverting the rate. We normalize the Lyapunov time by the maximum orbital period\footnote{The maximum of the three orbital periods corresponding to the three fundamental frequencies (see Section~\ref{sec:naff}).} of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
\begin{equation}
	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max} \, T_{\rm max}}
\end{equation}
We will use this as the prediction from the Lyapunov exponent for the timescale over which chaos should be dynamically important for a given orbit.

@@

Bounded, regular orbits in a triaxial potential have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion. The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}). \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses fast-Fourier transforms (FFTs) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$) to identify the frequencies. This method is referred to as `Numerical Approximation of Fundamental Frequencies' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems. We have implemented and tested a version of this procedure in the \project{Python} programming language. Our implementation differs slightly from the original definition and from that used in \cite{valluri98}; we refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}. For more details about the algorithm and differences with previous work, see \cite{laskar88, laskar93, papaphilippou96} and Appendix~\ref{sec:naffapdx}.

If an orbit is chaotic, the motion can no longer be expanded in terms of a single set of fundamental frequencies. For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time. \superfreq\ will pick out a set of frequencies for chaotic orbits that correspond to the largest peaks in the power spectrum of the orbits, however these peaks will change location and amplitude with time. For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two consecutive sections of orbit. The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window. We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $k$th fundamental frequency as
\begin{equation}
	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)} \, N} \label{eq:fdrate}
\end{equation}
where the upper index refers to the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit. By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
\begin{equation}
	t_\Omega = \, (\max_{a_k} \, \mathcal{R}_k)^{-1} \label{eq:fdtime}
\end{equation}
where the maximum is taken with respect to the corresponding amplitudes, $a_k$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.

@@

\todo{apw}{rename 'orbit structure' or 'orbit properties' and also show apocenter/pericenter map}
\subsection{Part I: Lyapunov and frequency diffusion times}\label{sec:results1}

We generate an isoenergy grid of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}). We fix $v_x = v_z = 0$, and compute $v_y$ from the energy. The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic. Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable numbers of stochastic tube orbits. The major classes of stable tube orbits circulate about either the major or minor axis. This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes. Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits. Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored either by the median pericentric distance (left) or median apocentric distance (right) after 16 orbital periods.

%Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the major orbit families are labeled. Points are colored (black / white) depending on whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods. Figure~\ref{fig:icorbits} show sample orbits from each of the main, labeled orbit families of the initial condition grid.

We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents. Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the logarithm of the Lyapunov time (in units of number of orbital periods) is mapped to greyscale intensity. The darker pixels have shorter Lyapunov times and are more chaotic. By fixing the integration time (in units of orbital periods, $\periods$), the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim 1000~\periods$).

For each orbit, we also separately integrate for $\approx$128 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 64 orbital periods. We have found that this is the shortest integration window for which we are able to successfully recover frequencies using \superfreq\ for $>$99.9\% of the orbits. Figure~\ref{fig:freqdiff} shows the same grid of initial conditions as in Figure~\ref{fig:lyapmap}, but now the greyscale intensity is set by the logarithm of the (order-unity) frequency diffusion time. The darker pixels have shorter frequency diffusion times and are more chaotic. This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent. The tube orbits in this potential are mostly regular or only mildly chaotic---the largest regular regions are associated with the short-axis and long-axis tube orbits---however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.

The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}). Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x,z) \approx(13,7)~{\rm kpc}$ and $(x,z) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}. The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 128 orbital periods.

While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods (thousands of Gigayears) and are thus not expected to be relevant for tidal stream evolution. In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to compare the effect of ordinary phase-mixing of tidal debris with potentially enhanced mixing due to chaos. We then compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.

@@

It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits. We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event. For a given set of orbital initial conditions---the `parent' orbit---we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time. 
\todo{apw}{We want to isolate phase mixing, so we want the ensembles to have the same spreads in frequencies...no variance introduced in the velocities}
The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}. If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
\begin{align}
	\delta x_{k,i} &\sim \mathcal{N}(0, r_{\rm tide}/\sqrt{3})\\
	\delta v_{k,i} &\sim \mathcal{N}(0, \sigma_v/\sqrt{3})\\
	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{M(<\|\bs{x}_0\|)}\right)^{1/3} \\
	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{M(<\|\bs{x}_0\|)}\right)^{1/3} 
\end{align}
where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm. These scalings are chosen so that the ensembles all start with approximately the same initial spread in frequencies (which also scales with $(m/M)^{1/3}$). We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.

We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular (non-resonant) orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S). The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, and the degree of chaos is taken from the \todo{apw}{nonresonant regular, weakly chaotic, strongly chaotic}
based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}. Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods. A thin `stream' forms on the regular orbit (left column), a more diffuse `stream' on the mildly chaotic orbit (middle column), and a `fanned' stream on the strongly chaotic orbit (right column). Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit ($t_\lambda \approx 55$, $t_\Omega \approx 8\times10^4$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream. \todo{apw}{numbers}

@@

\begin{table*}[ht]
\begin{center}
	\begin{tabular}{c | c c c c c c | c c }
		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
		R & 29.10 & 0 & 5.10 & 0 & 168.07 & 0 & $>10^3$ & $>10^8$\\
		\midrule
		W & 29.10 & 0 & 17.70 & 0 & 125.49 & 0 & $\approx55$ & $\approx8 \times 10^4$\\
		\midrule
		S & 29.10 & 0 & 22.10 & 0 & 99.45 & 0 & $\approx3.5$ & $\approx6 \times 10^3$\\
		\bottomrule
		\end{tabular}
	\caption{Orbits from the $xz$ grid with a range of chaotic timescales---R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit. Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods. Recall that the frequency diffusion time, $t_\Omega$, is the time over which we expect order-unity changes in the fundamental frequencies, hence why the timescales appear quite long. \label{tbl:three-orbits}}
\end{center}
\end{table*}

@@

Our main results and conclusions are summarized as follows:
\begin{enumerate}
	\item ``stream-fanning'' of tidal streams on weakly chaotic orbits \citep[as seen in simulations by][]{pearson15} occurs due to small-scale chaotic evolution of the fundamental frequencies of the debris star orbits;
	\item the Lyapunov time and frequency diffusion time are powerful indicators of chaos, but do not capture the importance of small-scale chaotic diffusion for the density evolution of small ensembles of orbits (tidal debris);
	\item tidal debris becomes diffuse and thus harder to observe on weakly chaotic orbits when the small-scale chaotic diffusion of the fundamental frequencies has a scale comparable to the internal spread in frequencies of the debris;
	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
	\item the variance of the fundamental frequencies of an orbit over a given time window is a strong predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
\end{enumerate}

@@

% Figure ??
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/ensembles.png}
\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}). The ensembles each contain 1000 particles and begin at pericenter rather than apocenter because they are meant to represent debris disrupted during a single (the first) pericentric passage. The particle positions qualitatively match the morphology of the `oldest' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations. } 
\label{fig:ensembles}
\end{center}
\end{figure}

% Figure ??
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/nbody.png}
\caption{ We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on the three orbits of Figure~\ref{fig:ensembles}. The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking. We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter. The mass of the progenitor is set to XX, and the length-scale of the Plummer sphere is set to 20~pc to get $\approx$10\% mass loss over the integration time. Panels show particle positions from the final snapshots of the simulations---for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis. These simulations confirm that the test-particle orbit ensembles do (qualitatively) capture the nature of the early-stripped tidal debris. } 
\label{fig:nbodysims}
\end{center}
\end{figure}

% Figure ??
\clearpage
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/ensemble-density.png}
\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}. The density distribution of the regular orbit ensemble stays compact and decreases slowly. The density of the weakly chaotic ensemble initially evolves faster to lower density and has a larger tail to lower densities. The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
\end{center}
\end{figure}

% Figure ??
\clearpage
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/ensemble-densities.png}
\caption{This is an alternate visualization of the density evolution of orbit ensembles around the three orbits of Figure~\ref{fig:ensemble-density}, showing only the evolution of the mean density for the orbits (normalized by the initial density of the ensemble). } 
\label{fig:densities}
\end{center}
\end{figure}

>>>>

An ensemble of regular orbits (e.g., tidal debris) will phase-mix due to (small) differences in the fundamental frequencies of the orbits. The frequency distributions of thin streams are generally close to one-dimensional---that is, one eigenvalue of the distribution of frequencies for an orbit ensemble will be much larger than the others because of the local shape of the Hessian of the potential \citep{sanders13, bovy14}. Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on the other eigenvalues of the frequency-space distribution.

Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and the chaotic evolution of the orbit and is thus difficult to predict. For example, ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume via the short-time chaotic diffusion processes, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}. In this work, we investigate the consequences of this short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}. We then discuss how this would affect our understanding of the coherence and density evolution of tidal streams.

@@

The MLE, $\lambda_{\rm max}$, is interpreted as a rate that quantifies the exponential divergence of infinitesimally close chaotic orbits. It is therefore useful to consider the corresponding $e$-folding time by inverting the rate,
\begin{equation}
	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max}}.
\end{equation}
We will use this as the prediction from the Lyapunov exponent for the timescale over which chaos should be dynamically important for a given orbit.

@@

Bounded, regular orbits in a triaxial potential have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion. The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}). \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses fast-Fourier transforms (FFTs) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$) to identify the frequencies. This method is referred to as `Numerical Approximation of Fundamental Frequencies' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems. We have implemented and tested a version of this procedure in the \project{Python} programming language. Our implementation differs slightly from the original definition and from that used in \cite{valluri98}; we refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub}.\footnote{\url{https://github.com/adrn/SuperFreq}} For more details about the algorithm and differences with previous work, see \cite{laskar88, laskar93, papaphilippou96} and Appendix~\ref{sec:naffapdx}.

If an orbit is chaotic, the motion can no longer be expanded in terms of a single set of fundamental frequencies. For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time. \superfreq\ will pick out a set of frequencies for chaotic orbits that correspond to the largest peaks in the power spectrum of the orbits, however these peaks will change location and amplitude with time. For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two consecutive sections of orbit. The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window. Following previous work, we define the fractional frequency diffusion rate, $\mathcal{R}$, in the $k$th fundamental frequency as
\begin{equation}
	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)}} \label{eq:fdrate}
\end{equation}
where the upper index refers to the two consecutive sections of orbit \citep{laskar93, valluri98, valluri12}. By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
\begin{equation}
	t_\Omega = \, (\max_{a_k} \, \mathcal{R}_k)^{-1} \label{eq:fdtime}
\end{equation}
where the maximum is taken with respect to the corresponding amplitudes, $a_k$, of the fundamental frequency components \citep[see][]{valluri12}.

@@

\subsection{Part I: Lyapunov and frequency diffusion times}\label{sec:results1}

We generate an isoenergy grid of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}). We fix $v_x = v_z = 0$, and compute $v_y$ from the energy. The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic. Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable numbers of stochastic tube orbits. The major classes of stable tube orbits circulate about either the major or minor axis. This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes. Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits. Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored by the median pericentric distance (left) and median apocentric distance (right) over an integration time of 64 orbital periods.

We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents. Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the color corresponds to the logarithm of the Lyapunov time. The darker pixels have shorter Lyapunov times and are more chaotic. Because of the fixed the integration time, the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $(t_\lambda/\periods) \gtrsim 1000$).

\todo{apw}{remove kpc units?}
For each orbit, we also separately integrate for $\approx$256 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 128 orbital periods. We have chosen this window size so that we recover the frequencies for regular orbits with fractional error $\approx10^{-8}$. With this integration window we are able to successfully recover frequencies using for $>$99.9\% of the orbits \superfreq. Figure~\ref{fig:freqdiff} shows the same grid of initial conditions as in Figure~\ref{fig:lyapmap}, but now the greyscale intensity is set by the logarithm of the frequency diffusion time. The darker pixels have shorter frequency diffusion times and are more chaotic. This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over just $\sim$100 orbital periods, compared to the many tens of thousands of orbits it would take to detect such features with the maximum Lyapunov exponent. The tube orbits in this potential are mostly regular or only mildly chaotic---the largest regular regions are associated with the short-axis and long-axis tube orbits---however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.

The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}). Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x_0,z_0) \approx(13,7)~{\rm kpc}$ and $(x_0,z_0) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}. The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 256 orbital periods.

While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods and are thus not expected to be relevant for tidal stream evolution. In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to compare the effect of ordinary phase-mixing of tidal debris with potentially enhanced mixing due to chaos. We then compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.

@@

It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits. We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event. For a given set of orbital initial conditions---the `parent' orbit---we integrate for 128 orbital periods, find the phase-space position of the minimum pericenter over this times, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time. We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits; all ensembles should start with the same initial spread in fundamental frequencies. We set the physical scale of the ensemble by the tidal radius in position and the velocity scale in velocity; the initial spread in fundamental frequencies will scale as $(m/M)^{1/3}$ like the tidal radius and velocity scale \citep[e.g.,][]{johnston98, apw14}. If $(\bs{x},\bs{v})$ are a set of initial conditions at pericenter for a parent orbit, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
\begin{align}
	\delta x_{k,i} &\sim \mathcal{N}(0, r_{\rm tide}/\sqrt{3})\\
	\delta v_{k,i} &\sim \mathcal{N}(0, \sigma_v/\sqrt{3})\\
	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{M(<\|\bs{x}_0\|)}\right)^{1/3} \\
	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{M(<\|\bs{x}_0\|)}\right)^{1/3} 
\end{align}
where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm. We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales. % we have verified this using regular orbits in a logarithmic potential...

We start by considering four particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D). The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, but have different frequency diffusion times; all four orbits are long-axis tube orbits. Figure~\ref{fig:orbits} shows the orbits in projection over a long integration period (1024 orbital periods). The initial conditions and chaos diagnostics for each orbit are listed in Table~\ref{tbl:orbit-info}. Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and evolved for 64 orbital periods. A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional--- stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with an interesting morphology (middle-right), and a `fanned' stream on the strongly chaotic orbit (right). Given the long Lyapunov and frequency diffusion times of the weakly chaotic orbit ($t_\lambda/\periods > 900$, $t_\Omega/\periods \approx 2\times10^5$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.

@@

\begin{table*}[ht]
\begin{center}
	\begin{tabular}{c | c c c c c c | c c }
		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda/\periods}$ & $\bs{t_\Omega/\periods}$ \\\toprule
A & 17 & 0 & 26.06 & 0 & 126.25 & 0 & $>$10^8 & XX\\
\midrule
B & 17 & 0 & 23.03 & 0 & 148.61 & 0 & XX & XX\\
\midrule
C & 17 & 0 & 25.35 & 0 & 131.68 & 0 & XX & XX\\
\midrule
D & 17 & 0 & 28.69 & 0 & 104.54 & 0 & XX & XX\\
		\bottomrule
		\end{tabular}
	\caption{Orbits from the $xz$ grid with a range of chaotic timescales---A is a regular, near-resonant orbit, B a regular, non-resonant orbit, C a weakly chaotic orbit, and S a strongly chaotic orbit. Initial conditions are given in physical units: Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$. Chaotic timescales are expressed ($t_\lambda$, $t_\Omega$) in number of orbital periods. Recall that the frequency diffusion time, $t_\Omega$, is the time over which we expect order-unity changes in the fundamental frequencies, hence why the timescales appear quite long. \label{tbl:orbit-info}}
\end{center}
\end{table*}

@@

Our main results and conclusions are summarized as follows:
\begin{enumerate}
	\item ``stream-fanning'' of tidal streams on weakly chaotic orbits \citep[as seen in simulations by][]{pearson15} occurs due to small-scale chaotic evolution of the fundamental frequencies of the debris star orbits;
	\item the Lyapunov time and frequency diffusion time are powerful indicators of chaos, but do not capture the importance of small-scale chaotic diffusion for the density evolution of small ensembles of orbits (tidal debris);
	\item tidal debris becomes diffuse and thus harder to observe on weakly chaotic orbits when the small-scale chaotic diffusion of the fundamental frequencies has a scale comparable to the internal spread in frequencies of the debris;
	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
	\item the variance of the fundamental frequencies of an orbit over a given time window is a strong predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
\end{enumerate}
\todo{apw}{make it clear: this is not about how much chaos, just timescale discrepancy}

@@

% Figure ??
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/four-orbits.png}
\caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D). All orbits are long-axis tubes with similar pericenters and apocenters. Orbits in this Figure were integrated for 1024 orbital periods and are shown in projection.} 
\label{fig:orbits}
\end{center}
\end{figure}

% Figure ??
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/ensembles.png}
\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}). The ensembles each contain 1024 particles and are initialized at pericenter. The particle positions qualitatively match the morphology of the `oldest' debris from the corresponding N-body simulations (Figure~\ref{fig:nbodysims}). } 
\label{fig:ensembles}
\end{center}
\end{figure}

% Figure ??
\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/nbody.png}
\caption{ We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on the three orbits of Figure~\ref{fig:ensembles}. The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking. We start the progenitor systems at apocenter and evolve for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter. The mass of the progenitor is set to $10^4~\msun$, and the length-scale of the Plummer sphere is set to 5~pc to get $\approx$50--75\% mass loss over the integration time. Panels show particle positions from the final snapshots of the simulations---for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis. These simulations confirm that the test-particle orbit ensembles (Figure~\ref{fig:ensembles}) do (qualitatively) capture the nature of the early-stripped tidal debris. } 
\label{fig:nbodysims}
\end{center}
\end{figure}

% Figure ??
%\clearpage
%\begin{figure}[p]
%\begin{center}
%\includegraphics[width=\textwidth]{figures/ensemble-density.png}
%\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}. The density distribution of the regular orbit ensemble stays compact and decreases slowly. The density of the weakly chaotic ensemble initially evolves faster to lower density and has a larger tail to lower densities. The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
%\end{center}
%\end{figure}
